---
title: "Lab3_CatieKitrinos"
output: html_document
---

```{r}
library(tidyverse)

```
#####Inspecting Data Frame Objects
```{r}
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data_raw/portal_data_joined.csv")
```


```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv")
```
```{r}
str(read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = TRUE))
```

surveys
```{r}
str(surveys)
```

```{r}
head(surveys)
```
```{r}
view(surveys)
```

## Challenge1 : Based on the output of str(surveys), can you answer the following questions?
#1: What is the class of the object surveys?
#2: How many rows and how many columns are in this object?
#3: How many species have been recorded during these surveys?

```{r}
str(surveys)
```

## Answers:
#class: data frame
#rows: 34786. columns: 13
#species: 48

## Indexing and Subsetting Data Frames

# first element in the first column of the data frame (as a vector)
surveys[1, 1]   
# first element in the 6th column (as a vector)
surveys[1, 6]   
# first column of the data frame (as a vector)
surveys[, 1]    
# first column of the data frame (as a data.frame)
surveys[1]      
# first three elements in the 7th column (as a vector)
surveys[1:3, 7] 
# the 3rd row of the data frame (as a data.frame)
surveys[3, ]    
# equivalent to head_surveys <- head(surveys)
head_surveys <- surveys[1:6, ] 

#surveys[, -1]          
#The whole data frame, except the first column
#surveys[-(7:34786), ] 
#Equivalent to head(surveys)
#surveys["species_id"]       # Result is a data.frame
#surveys[, "species_id"]     # Result is a vector
#surveys[["species_id"]]     # Result is a vector
#surveys$species_id          # Result is a vector


## Challenge2: 

#1. Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.
#2.  Notice how nrow() gave you the number of rows in a data.frame?
  #Use that number to pull out just that last row in the data frame.
  #Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.
  #Pull out that last row using nrow() instead of the row number.
  #Create a new data frame (surveys_last) from that last row.
#3. Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.
#4. Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.

## 1Answer
```{r}
surveys_200 <-surveys[200,]

```

```{r}
surveys_200
```

## 2 Answer
```{r}
surveys[34786,]
```
```{r}
tail(surveys)
```

```{r}
n_rows <-nrow(surveys)
```
```{r}
surveys_last <-surveys[n_rows, ]
```
```{r}
surveys_last
```

## 3 Answer
```{r}
surveys_middle <- surveys[n_rows / 2, ]
```

surveys_middle

## 4 Answer

```{r}
surveys_head <- surveys[-(7:n_rows), ]
surveys_head
```
## Factors

```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```

sex # current order

```{r}
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```

as.character(sex)

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(levels(year_fct))[year_fct]
```

#bar plot of the number of females and males captured during the experiment:
```{r}
sex <- factor(surveys$sex)
head(sex)
levels(sex)
levels(sex)[1] <- "undetermined"
levels(sex)
head(sex)

```

## Challenge 3: Rename “F” and “M” to “female” and “male” respectively.Now that we have renamed the factor level to “undetermined”, can you recreate the barplot such that “undetermined” is last (after “male”)?


# Answer1
```{r}
levels(sex)[2:3] <- c("female", "male")
levels(sex)
head(sex)
```

## Answer2
```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```

## UsingStringsasFactors=False

#Compare the difference between our data read as `factor` vs `character`.

```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```

#Convert the column "plot_type" into a factor
```{r}
surveys$plot_type <- factor(surveys$plot_type)
```

## Challenge 4: 

#We have seen how data frames are created when using read.csv(), but they can also be created by hand with the data.frame() function. There are a few mistakes in this hand-crafted data.frame. Can you spot and fix them? Don’t hesitate to experiment!

#Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):
#Are they what you expected? Why? Why not?
#What would have been different if we had added stringsAsFactors = FALSE when creating the data frame?
#What would you need to change to ensure that each column had the accurate data type?

## Answer1
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "squishy", "spiny", "poofy"),
          weight = c(45, 8, 1.1, 0.8)
          )

animal_data
```

#1a: missing quotations for characters, missing comma after 8 in weight category, and missing a 4h variable in the "feel category.

## Answer2
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
str(country_climate)
```


#2a:Yes, I can see the class (factor=country,climate, temp, northernhemisphere, number=haskangaroo)
#2b:They are not what I would expect because temperatiure is not a number  and has kangaroo is a number
#2c: Setting stringsAsFactors= FALSE would not make them factors but vectors
#2d: Taking out the quotes from "15" andf "FALSE" then making 1 a "TRUE" would make the classes correct


## Formatting Dates

```{r}
str(surveys)
library(lubridate)
my_date <- ymd("2015-01-01")
str(my_date)
```

# sep indicates the character to use to separate each component
```{r}
my_date <- ymd(paste("2015", "1", "1", sep = "-")) 
str(my_date)
paste(surveys$year, surveys$month, surveys$day, sep = "-")
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))

```

```{r}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
str(surveys)
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]
head(missing_dates)

```
## Quenstion 5: Why did these dates fail to parse? If you had to use these data for your analyses, how would you deal with this situation?

#Answer : They failed to parse because the "day" is listed as 31 for all of these when these months only have 30 days
