---
title: "Lab4_CatieKitrinos"
output: html_document
---
#load tidyverse
```{r}
library(tidyverse)
```
#uploaddataset
```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
#checkstructure
```{r}
str(report_03_11_2020)
```
#Viewdata
```{r eval=FALSE, include=FALSE}
View(report_03_11_2020)
```
#installdt
```{r}
library(DT)
datatable(report_03_11_2020)
```

## Exercise: 1 

## Part 1: The format of the daily reports has changed over time. What are the differences between 3/13/2020, the Friday before spring break 2020 and yesterday 6 months later on 9/13/2020? Load the files into R and use str() and View() (or in the top right corner click on Environment and then the data frame.). Note the in the more recent dates they have changed the column headers so that it is no longer necessary to rename Province_State and Country_Region.

# Upload 3/13 File
```{r}
report_03_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```
#View 3/13
```{r eval=FALSE, include=FALSE}
view(report_03_13_2020)
```
#See strcuture 3/13
```{r}
str(report_03_13_2020, stringsAsFactors = TRUE)
```
#Count Countries

```{r}
unique(report_03_13_2020$Country_Region)
```
#Upload 9/13 File

```{r}
report_09_13_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv"))
```
#View 9/13 File
```{r eval=FALSE, include=FALSE}
view(report_09_13_2020)
```
#see structure of 9/13
```{r}
str(report_09_13_2020, stringsAsFactors = TRUE)
```
#Count Countries
```{r}
unique(report_09_13_2020$Country_Region)
```


## Answer Part 1: There are now 14 rather than 8 columns where (FIPS, Admin, Recovered, Active, Combined Key, Incidence Rate, and Case fatality are all new). There were 125 countries on the list in 3/13 now there are 187

## Part 2: Many countries have multiple entries representing different regions (e.g. counties and states). To get the counts for a country we need to summarise the data. Use group_by() and summarize() to make a data frame (tibble) with total counts for each country.

## Answer Part 2: 
```{r}
 country_counts <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarize(cases = sum(Confirmed))
```
```{r}
 country_counts
```

## Part 3: To get counts for confirmed cases per U.S state use filter() to select US and group_by and summarize as above.

## Answer Part 3: 

```{r}
 state_count <- report_09_13_2020 %>%
  filter(Country_Region == "US") %>%
  group_by(Province_State) %>%
  summarize(cases = sum(Confirmed))

state_count
```
## Part 4: It will be difficult to make simple graphs with the data from all countries in the world or even the US states. Use arrange_by() and slice() to get just the top 10 countries for Deaths.

## Answer Part 4:
```{r}
death_counts <- report_09_13_2020 %>%
  group_by(Country_Region) %>%
  summarize(deaths = sum(Deaths)) %>%
  arrange(desc(deaths))

death_counts

  top_10_death <- slice(death_counts, 1:10)
  
  top_10_death
```



