---
title: "Lab5_CatieKitrinosMarkdown"
output: html_document
---

```{r}

library("tidyverse")

```


```{r load-packages, include=FALSE}
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
```


```{r}

surveys_complete <- read_csv("data_raw/surveys_complete.csv")

```


# Plotting time series data

```{r}
yearly_counts <- surveys_complete %>%
  count(year, genus)
```


```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
     geom_line()
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, group = genus)) +
    geom_line()
```

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, color = genus)) +
    geom_line()
```

# Integrating the pipe operator with ggplot2

```{r}
yearly_counts %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```

```{r}
yearly_counts_graph <- surveys_complete %>%
    count(year, genus) %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()

yearly_counts_graph

```

# Faceting

```{r}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
    geom_line() +
    facet_wrap(facets = vars(genus))
```

## Split the line by sex

```{r}
 yearly_sex_counts <- surveys_complete %>%
                      count(year, genus, sex)

 yearly_sex_counts
```
```{r}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(facets =  vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```

### One column, facet by rows

```{r}

ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```

## One row, facet by column

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols = vars(genus))
```

```{r}
 ggplot(data = yearly_sex_counts, 
        mapping = aes(x = year, y = n, color = sex)) +
     geom_line() +
     facet_wrap(vars(genus)) +
     theme_bw()
```

## Challenge: Use what you just learned to create a plot that depicts how the average weight of each species changes through the years.

```{r}
yearly_weight <- surveys_complete %>%
                group_by(year, species_id) %>%
                 summarize(avg_weight = mean(weight))
```
```{r}
ggplot(data = yearly_weight, mapping = aes(x=year, y=avg_weight)) +
   geom_line() +
   facet_wrap(vars(species_id)) +
   theme_bw()
```

## Customization

```{r}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals") +
    theme_bw()
```

```{r fig.height = 10}

ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(text=element_text(size = 16))

```

```{r fig.height= 10}

ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))

```

```{r}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, 
                                               vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot() +
    grey_theme
```

## Challenge: Challenge: With all of this information in hand, please take another five minutes to either improve one of the plots generated in this exercise or create a beautiful graph of your own. Use the RStudio ggplot2 cheat sheet for inspiration.

##Here are some ideas:

##See if you can change the thickness of the lines.
##Can you find a way to change the name of the legend? What about its labels?
##Try using a different color palette (see ##http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/).


```{r fig.height = 4}

ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line( size = 1.5) +
    facet_wrap(vars(genus)) +
    theme_dark() +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals"
        )
  
```



## Arranging and exporting plots

```{r fig.height= 10}

library(gridExtra)

spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             aes(x = species_id, y = weight)) +
  geom_boxplot() +
  labs(x = "Species", 
       y = expression(log[10](Weight))) +
  scale_y_log10() + 
  labs()

spp_count_plot <- ggplot(data = yearly_counts, 
                         aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))

spp_count_plot

```

##Saving in High Resolution

```{r}
my_plot <- ggplot(data = yearly_sex_counts, 
                  aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 12),
          text = element_text(size = 16))

ggsave("name_of_file.png", my_plot, width = 15, height = 10)

## This also works for grid.arrange() plots
combo_plot <- grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, 
                           widths = c(4, 6))
ggsave("combo_plot_abun_weight.png", combo_plot, width = 10, dpi = 300)

```


```{r}
head(surveys_complete)
```
















