---
title: "Lab5b_CatieKitrinosMarkup"
output: html_document
---

# Joining Tables

```{r}
library(tidyverse)
```

```{r}
Confirmed_State_3_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r}
str(Confirmed_State_3_13)
```

```{r}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r}
str(Confirmed_State_9_13)
```

```{r}
setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_3_13$Province_State)

```

```{r}
Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```

```{r}
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_3_13_9_13_joined)
```
```{r}
tail(Confirmed_State_3_13_9_13_joined, 5)
```

```{r}
which(is.na(Confirmed_State_3_13_9_13_joined))
```

```{r}
 Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_3_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_3_13_2020 = 0))
head(Confirmed_State_3_13_9_13_joined)


```

# Switching between wide and long table formats.

```{r}
Confirmed_State_3_13_9_13_joined_long <- Confirmed_State_3_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```

## to adjust figure size {r, fig.width = 5, fig.height = 10}

```{r}
ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + 
    geom_point(aes(color = Date)) 
```

# Working with the time series data

```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
head(time_series_confirmed)
```
```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 
```

```{r}
head(time_series_confirmed_long)
```
# Import Deaths

```{r}
 download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "data_raw/time_series_covid19_deaths_global.csv")
```

```{r}
time_series_deaths <- read_csv("data_raw/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths
```

```{r}
time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)
```
# Joining the time series tables

```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)
```

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```

```{r}
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)
```
```{r}
which(is.na(time_series_long_joined$Confirmed))
```

```{r}
which(is.na(time_series_long_joined$Deaths))
```

```{r}
library(lubridate)
```

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```

# Making Graphs from the time series data

##To make a times series graph of the confirmed cases we need to summarize the Country date to count up the individual state data for the US.

```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")
```


```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths") +
    facet_wrap(~Country_Region, ncol=2, scales="free_y")
```


```{r}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
```


```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")
```


# Exercises: 

## 1.Finish the Data Visualization tutorial and add it your GitPage web site
#####(see 5a in github repository)

## 2.Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height in the R code chunk header.

##Import 6/13 data set

```{r}
Confirmed_State_6_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/06-13-2020.csv")) %>%
  rename(Country_Region = "Country_Region", Province_State = "Province_State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

##Check structure

```{r}
str(Confirmed_State_6_13)
```

##Joining

```{r}
Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_6_13_9_13_joined)
```
##Read tail

```{r}
tail(Confirmed_State_3_13_9_13_joined, 5)
```

##Prune NAs

```{r}
which(is.na(Confirmed_State_6_13_9_13_joined))
```

##Plot

```{r}
Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_6_13_2020 = 0))
head(Confirmed_State_6_13_9_13_joined)

```
```{r}
Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```

##Create bar plot

```{r fig.height=10, fig.width=15}
rainbow_bar <- ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State, color = Province_State))  + 
    geom_bar(stat = 'identity') 
rainbow_bar

```


## 3.Add more informative x and y labels and a title to the above graph in Ex2.

```{r fig.height=10, fig.width=15}
rainbow_bar2 <- rainbow_bar <- ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State, color = Province_State))  + 
      labs(title = "Covid 19 Confirmed Cases 6/13 & 9/13 by US State",
         x = "Number of Confirmed Cases",
         y = "State") +
    geom_bar(stat = 'identity') 

rainbow_bar2
```



## 4.Using the time series data. Make a plot of the total number of confirmed deaths per day worldwide

##deaths file upload
```{r}
time_series_deaths
```
##Pivot

```{r}
time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)
```
##Create Daily Counts

```{r}
time_series_deaths_long %>%
  group_by(Date) %>%
  summarize(daily_death = sum(Deaths, na.rm = TRUE))

```

##Plot 

```{r fig.height= 7}

time_series_deaths_long %>%
  group_by(Date) %>%
  summarize(daily_death = sum(Deaths)) %>%
ggplot(aes(x = Date,  y = daily_death)) + 
    geom_point( size = 3) +
    geom_line( size = 3) +
    ggtitle("COVID-19 Worldwide Deaths")

```



## 5.Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

# check time series joined

```{r}
time_series_long_joined_counts
```


```{r}
worldwide_counts <- time_series_long_joined_counts %>% 
  group_by(Date) %>% 
  mutate(daily_count = sum(Counts))
  
worldwide_counts
```


## 6.Plot US deaths/confirmed cases per day.

```{r}
worldwide_counts %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = daily_count)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Total Cases / Deaths")
```


## 7.Make a single graph with from the times series data with the 10 countries with the hightest death total.


```{r}
arr_deaths <- time_series_deaths_long %>% 
    group_by(Country_Region) %>% 
    summarise_at(c("Deaths"), sum) %>%
    arrange(desc(Deaths))
    
arr_deaths %>%
    slice(Deaths, 1:10) %>%
    arrange(desc(Deaths)) %>%
    slice(Deaths, 1:10)

```

```{r fig.height = 5}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("US", "Brazil","United Kingdom","Italy","Mexico", "France","Spain","India","Iran","Peru")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")

```
## 8.Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

```{r fig.height = 6}

facet_graph <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Deaths"), sum) %>% 
    filter (Country_Region %in% c("US", "Brazil","United Kingdom","Italy","Mexico", "France","Spain","India","Iran","Peru")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point( size = .3) +
    geom_line( size = .2) +
    ggtitle("COVID-19 Deaths") +
    theme(text=element_text(size = 5)) +
    facet_wrap(~Country_Region, ncol=2, scales="free_y")
    
```


```{r}

facet_graph

```


## 9.Using Facet_Wraps make a graph from the times series data for confirmed cases for all US states and territories. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

# upload
```{r}
US_time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"))

head(US_time_series_confirmed)
```

```{r}
view(US_time_series_confirmed)
```

# pivot

```{r}

US_time_series_confirmed_long <- US_time_series_confirmed %>%
  pivot_longer(-c(UID, iso2,iso3, code3, FIPS, Admin2, Lat, Long_, Combined_Key, Province_State, Country_Region, Lat, Long_),
               names_to = "Date", values_to = "Confirmed")

head(US_time_series_confirmed_long)

```

```{r fig.height=10, fig.width=15}

facet_graph_US <- US_time_series_confirmed_long %>% 
    group_by(Province_State, Date) %>% 
    summarise_at(c("Confirmed"), sum) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Province_State)) + 
    geom_point( size = .3) +
    geom_line( size = .3) +
    ggtitle("COVID-19 US State") +
    facet_wrap(~Province_State, ncol=5, scales="free_y")

facet_graph_US

```

## 10.Apply a theme other than the one used in the Data Carpentries tutorial to the graph in Ex8.

```{r fig.height = 8}


facet_graph_dark <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Deaths"), sum) %>% 
    filter (Country_Region %in% c("US", "Brazil","United Kingdom","Italy","Mexico", "France","Spain","India","Iran","Peru")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths") +
    theme_dark() + 
    facet_wrap(~Country_Region, ncol=2, scales="free_y") 
    
facet_graph_dark

```
















































