---
title: "Lab9_CKitrinosMarkdown"
output: html_document
---

# Matching Patterns with Regular Expressions

```{r}
library(stringr)
library(tidyverse)
```

# Basic Matches

```{r}
x <- c("apple", "banana", "pear")

str_view(x, "an")
```

```{r}
str_view(x, ".a.")

```

```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```

```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

# Exercise 1: 

##1: "/" Doesn't match / because this is an escape character. "//" does not match / because this only escapes the escape character and creates a string. "///" doesnt match / because it is matching the escape character.

# Anchors

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

# Exercise 14.3.2.1

##1: You would need to use str_view("^\\$\\^\\$$") to escape the function of the characters and create a string

# Character Classes and Alternatives

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

#Exercise 3

##1:

```{r}
str_subset(stringr::words, "^[aeiou]")
str_subset(stringr::words, "[aeiou]", negate=TRUE)
str_subset(stringr::words, "[^e]ed$")
str_subset(stringr::words, "i(ng|se)$")
```

# Repetition

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, 'C{2,3}?')
```

```{r}
str_view(x, 'C[LX]+?')
```

# 14.3.4.1 Exercises

##1: {0,1} match 1 at maximum , {1,} mostgreater or equal to 1, {0,} match greater or equal to 0

# 14.3.5 Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# 14.3.5.1 (.)\1\
##1: The same character appearing three times in a row. E.g. "aaa"
"(.)(.)\\2\\1": A pair of characters followed by the same pair of characters in reversed order. E.g. "abba".
(..)\1: Any two characters repeated. E.g. "a1a1".
"(.).\\1.\\1": A character followed by any character, the original character, any other character, the original character again. E.g. "abaca", "b8b.b".
"(.)(.)(.).*\\3\\2\\1" Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order.

#14.4 Tools

##14.4.1 Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")

```

```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

## Use filter instead

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```

# 14.4.1.1 Exercises

```{r}
words[str_detect(words, "^x|x$")]
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```

## Extract sentences


```{r}
length(sentences)
```

```{r}
head(sentences)
```

## Select sentences that contain a color

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
has_colour
```

## First select sentences with more than 1 match

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
str_extract_all(more, colour_match)
```

## Create matrices of short matches

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

##14.4.2.1 Exercises

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)

```

## 14.4.3 Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>% 
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

## 14.4.3.1 Exercises

##1:

```{r}
nuword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, nuword)] %>%
  str_extract(nuword)
```

## 14.4.4 Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

## Flip the orders of the second and third words

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

## 14.4.4.1 Exercises

```{r}
str_replace_all("past/present/future", "/", "\\\\")
#> [1] "past\\present\\future"

```

## 14.4.5 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

## Extract first number of the list

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

## 14.4.5.1 Exercises

```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]

```

## 14.7 stringi

##NA



















