---
title: "Lab9x_CKitrinosMarkdown"
output: html_document
---

# Control Structures

## Conditional Statements with 'if-else'

if(<condition>) {
        ## do something
} 
## Continue with rest of code

## Add else clause

if(<condition>) {
        ## do something
} 
else {
        ## do something else
}

## Some tests

if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}

## example

## Generate a uniform random number
x <- runif(1, 0, 10)  
if(x > 3) {
        y <- 10
} else {
        y <- 0
}
# My example

```{r}
## Generate a uniform random number
x <- runif(2, 3, 5)  
if(x < 5) {
        y <- 5
} else {
        y <- 0
}
```


# For loops

```{r}
for(i in 1:10) {
        print(i)
}
```

```{r}
x <- c("a", "b", "c", "d")
for(i in 1:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

# My for loop

```{r}
x <- c("lemur", "capuchin", "baboon", "tarsier")
for(i in 1:4) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```

```{r}
for(letter in x) {
        print(letter)
}
```

```{r}
for(i in 1:4) print(x[i])
```

# Nested for loops

```{r}
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```

# While loops

```{r}
count <- 0
while(count < 5) {
        print(count)
        count <- count + 1
}
```

```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```

# Functions

```{r}
f <- function() {
        ## This is an empty function
}
## Functions have their own class
class(f)  
```

```{r}
## Execute this function
f()       
```

```{r}
f <- function() {
        cat("Its freezing\n")
}
f()
```

```{r}
f <- function(num) {
        for(i in seq_len(num)) {
                cat("Its freezing\n")
        }
}
f(3)
```

```{r}
f <- function(num) {
        hello <- "John you scrooge turn the heat on\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(3)
```

```{r}
print(meaningoflife)
```

```{r eval=FALSE, include=FALSE}
f()
```

```{r}
f <- function(num = 1) {
        hello <- "You'll be sorry when the pipes freeze\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
f()    ## Use default value for 'num'
```

```{r}
f(4)   ## Use user-specified value
```

```{r}
f(num = 5)
```


























